Dryad Data Package for study "DateLife: leveraging databases and analytical tools to reveal the dated Tree of Life

### 2022-07-09 dryad/README.Rmd
formatting description of csv files

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  results = 'hide'
)
```

@@

```{r, echo = FALSE}
text <- c()
for (file_name in csv_file_names) {
  # read the csv file in:
  tt <- utils::read.csv(file = paste0("../dryad/", file_name), header = TRUE, row.names = NULL)
  # remove the column with row numbers:
  tt <- tt[,-1]
  # create the description text
  var_list <- paste("\\tab", gsub("[.]", " ", colnames(tt)))
  var_list <- paste(var_list, collapse = ": \\newline")
  text <- paste(text, 
                paste("DATA-SPECIFIC INFORMATION FOR:", file_name, "\\newline"),
                "1. Number of variables: ",
                ncol(tt), "\\newline",
                "2. Number of cases/rows:",
                nrow(tt), "\\newline",
                "3. Variable List: \\newline", var_list, "\\newline",
                "4. Missing data codes: None \\newline
                5. Abbreviations used: NA = not applicable \\newline \\newline")
}
```

`r text`
DATA-SPECIFIC INFORMATION FOR: Rye_2021_a_Cohort_IDs.xlsx

1. Number of variables: 4

2. Number of cases/rows: 73

3. Variable List:
	Screening_cohort: Patient included in screening cohort; yes/no
	Screening_cohort_ID: Patient ID number for screening cohort
	Site_inv_cohort: Patient included in site investigation cohort; yes/no
	Site_inv_cohort_ID: Patient ID number for site investigation cohort

4. Missing data codes:
	None

5. Abbreviations used:
	N/A; not applicable

6. Other relevant information:
	20 of 21 patients in the screening cohort whose tumour tested positive for F. nucleatum, B. fragilis or both species were included in the site investigation cohort, along with 31 additional patients. Tumour material for one screening cohort patient whose tumour was positive for both species (patient 24) was not available for further DNA extraction.

DATA-SPECIFIC INFORMATION FOR: Rye_2021_b_Screening_qPCR_data.xlsx

1. Number of variables: 9

2. Number of cases/rows: 42

3. Variable List:
	Screening_cohort_ID: Screening cohort patient ID number
	B-actin_mean_Ct: Mean Ct for beta-actin; no. cycles
	PGT_mean_Ct: Mean Ct for prostaglandin transporter; no. cycles
	F_nucleatum_mean_Ct: Mean Ct for Fusobaterium nucleatum; no. cycles
	B_fragilis_gyrase_mean_Ct: Mean Ct for Bacteroides fragilis gyrase; no. cycles
	B_fragilis_bft_mean_Ct: Mean Ct for Bacteroides fragilis toxin; no. cycles
	B_breve_mean_Ct: Mean Ct for Bifidobacterium breve; no. cycles
	C_showae_mean_Ct: Mean Ct for Campylobacter showae; no. cycles
	L_buccalis_mean_Ct: Mean Ct for Leptotrichia buccalis ; no. cycles

4. Missing data codes:
	No Amp; No amplification

5. Abbreviations used:
	Amp; amplification, BFT; Bacteroides fragilis toxin, PGT; prostaglandin transporter

6. Other relevant information:
	Reactions were performed in duplicate. Results were reported where one or both samples amplified. Values represent the mean Ct where both samples amplified or the individual Ct for single amplifications.

DATA-SPECIFIC INFORMATION FOR: Rye_2021_c_Site_inv_species_by_site.xlsx

1. Number of variables: 9 + comments field

2. Number of cases/rows: 436

3. Variable List:
	Site_inv_cohort_ID: Site investigation cohort patient ID number
	Tumour_ID: Tumour ID number for patients with synchronous tumours (n = 5); 1/2
	Site: Area of tissue targeted; Normal (proximal), Normal (distal), Normal (adjacent), Tum luminal surface, Tumour, Tumour (mutinous), Invading margin, Inflammation, Stroma, Lymph node
	Site_code: Numerical code assigned to site; 1-11
	Site_code_2: Second site if same area used to represent more than one site; 1-11
	B_actin_mean_Ct: Mean Ct for beta-actin; no. cycles
	PGT_mean_Ct: Mean Ct for prostaglandin transporter; no. cycles
	F_nucleatum_mean_Ct: Mean Ct for Fusobaterium nucleatum; no. cycles
	B_fragilis_mean_Ct: Mean Ct for Bacteroides fragilis gyrase; no. cycles
	Comments: Used to flag values excluded from analysis

4. Missing data codes:
	No Amp; No amplification
	Single Amp; Single amplification
	N/A: not applicable

5. Abbreviations used:
 	Amp; amplification, PGT; prostaglandin transporter, Tum; tumour

6. Other relevant information:
	Target sites included: normal tissue from the proximal and distal surgical margins of the resection specimen, normal tissue adjacent to the tumour region, the tumour luminal surface, central tumour, mucinous tumour (where applicable), invading margin, sites of inflammation, stroma, lymph nodes containing tumour deposits and metastatic sites where available. Metastatic site data have been excluded from this dataset to minimise any risk of deductive disclosure of patient identities.
	B-actin and PGT reactions were performed in duplicate and all samples amplified.
	F. nucleatum and B. fragilis reactions were performed in triplicate and results reported where two or more samples amplified.  Mean cT values were excluded from analysis if the SD of the replicate Ct values > 5 (highlighted red).

DATA-SPECIFIC INFORMATION FOR: Rye_2021_d_Site_inv_16S_qPCR_data.xlsx

1. Number of variables: 4

2. Number of cases/rows: 51

3. Variable List:
	Site_inv cohort_ ID: Site investigation cohort patient ID number
	PGT_mean_Ct: Mean Ct for prostaglandin transporter; no. cycles
	16S_TFS_mean_Ct: Mean Ct for 16S rRNA using primer set 1
	16S_IDT_mean_Ct: Mean Ct for 16S rRNA using primer set 2

4. Missing data codes:
	None

5. Abbreviations used:
	IDT; Integrated DNA Technologies, PGT; prostaglandin transporter, TFS; Thermo Fisher Scientific

6. Other relevant information:
	Primer set 1: Obtained from Integrated DNA Technologies; Sequences as published in Nadkarni et al. Determination of bacterial load by real-time PCR using a broad-range (universal) probe and primers set. Microbiology. 2002;148(Pt 1):257-66. doi: 10.1099/00221287-148-1-257. PubMed PMID: 11782518.
	Primer set 2: Obtained from Thermo Fisher Scientific; Assay ID Ba04230899_s1.
	Performed on DNA samples extracted from the tumour luminal surface.

DATA-SPECIFIC INFORMATION FOR: Rye_2021_e_Seq_data_absolute_abundance.xlsx

1. Number of variables: 7

2. Number of cases/rows: 308

3. Variable List:
	OTU_ID: OTU ID number
	Pt_32: Absolute abundance of each OTU for patient 32; no. sequence reads
	Pt_26: Absolute abundance of each OTU for patient 26; no. sequence reads
	Pt_35: Absolute abundance of each OTU for patient 35; no. sequence reads
	Pt_28: Absolute abundance of each OTU for patient 28; no. sequence reads
	Pt_29: Absolute abundance of each OTU for patient 29; no. sequence reads
	Consensus_Lineage: Consensus lineage

4. Missing data codes:
	None

5. Abbreviations used:
	OTU; Operational Taxonomic Unit, Pt; patient

6. Other relevant information:
	Performed on DNA samples extracted from the tumour luminal surface.
	Taxonomy assigned using the QIIME 1 default classifier, pre-trained against Greengenes database5 (Version 13_8, Aug 2013).

DATA-SPECIFIC INFORMATION FOR: Rye_2021_f_Seq_data_otu_table.xlsx

1. Number of variables: 6

2. Number of cases/rows:
	Sheet 1 'Phylum_level': 13
	Sheet 2 'Class_level': 21
	Sheet 3 'Order_level': 28
	Sheet 4 'Family_level': 49
	Sheet 5 'Genus_level': 85
	Sheet 6 'Species_level': 101

3. Variable List:
	Taxon: Assigned taxon
	Pt_32: Relative abundance of each taxon for patient 32; proportion of total sequence reads
	Pt_26: Relative abundance of each taxon for patient 26; proportion of total sequence reads
	Pt_35: Relative abundance of each taxon for patient 35; proportion of total sequence reads
	Pt_28: Relative abundance of each taxon for patient 28; proportion of total sequence reads
	Pt_29: Relative abundance of each taxon for patient 29; proportion of total sequence reads

4. Missing data codes:
	None

5. Abbreviations used:
	Pt; patient

6. Other relevant information:
	Performed on DNA samples extracted from the tumour luminal surface.
	Taxonomy assigned using the QIIME 1 default classifier, pre-trained against Greengenes database5 (Version 13_8, Aug 2013).

DATA-SPECIFIC INFORMATION FOR: Rye_2021_g_Seq_data_mg_blast.xlsx

1. Number of variables: 16

2. Number of cases/rows: 366

3. Variable List:
	subject: OTU ID
	gi: Nucleotide BLAST accession number
	evalue: Expectation Value
	bit-score: Bit score
	score: Raw alignment score
	alignment-length: Alignment length; number of bases
	%identity: Percentage of bases identical to OTU group reference sequence
	identical: Number of bases identical to OTU group reference sequence
	positives: Number of bases with a positive match to OTU group reference sequence
	%positives: Percentage of bases with a positive match to OTU group reference sequence
	scientific-names: Scientific name of OTU classification group
	common-names: Common name of OTU classification group
	subject-blast-names: Grouping name
	subject-super-kingdoms: Kingdom names
	subject-title: Descriptive title for OTU group
	OTU-seq: OTU reference Sequence

4. Missing data codes:
	None

5. Abbreviations used:
	OTU; Operational Taxonomic Unit

6. Other relevant information:
	Taxonomy assigned using the QIIME 1 default classifier, pre-trained against Greengenes database5 (Version 13_8, Aug 2013).

>>>>

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE,
  results = 'hide'
)
```

@@

```{r}

var_description_1 <- c("Node names in tree topology, as defined by taxon A and taxon B",
                       "Scientific name of the first element of a taxon pair that defines a node from a tree topology",
                       "Scientific name of the second element of a taxon pair that defines a node from a tree topology",
                       "Ages of nodes defined by taxon A and taxon B, in Million years",
                       "Reference of the study that published the chronogram from which the node ages were extracted")

var_description_2 <- c("Names of nodes from the tree topology used for the dating analysis",
                       "Minimum node age, in Million years",
                       "First quartile of node age distributions, in Million years",
                       "Mean node age, in Million years",
                       "Median node age, in Million years",
                       "Third quartile of node age distributions, in Million years",
                       "Maximum node age, in Million years",
                       "Variance of node age, in Million years",
                       "Standar deviation of node age, in Million years")
```

```{r}
text <- c()
var_descriptions <- list(var_description_1, var_description_2, var_description_1, var_description_2) 
names(var_descriptions) <- csv_file_names

for (file_name in csv_file_names) {
  # read the csv file in:
  tt <- utils::read.csv(file = paste0("../dryad/", file_name), header = TRUE, row.names = NULL)
  # remove the column with row numbers:
  tt <- tt[,-1]
  # create the description text
  var_list <- gsub("[.]", " ", colnames(tt))
  var_list <- paste0("\\tab ", var_list, ": ", var_descriptions[[file_name]], "\\n ")
  var_list <- paste0(var_list, collapse = "")
  print(var_list)
  text <- paste(text, 
                paste("DATA-SPECIFIC INFORMATION FOR:", file_name, "\\n "),
                "\\n1. Number of variables/columns:", ncol(tt), "\\n ",
                "\\n2. Number of cases/rows:", nrow(tt), "\\n ",
                "\\n3. Variable List \\n ", var_list,
                "\\n4. Missing data codes: None \\n ",
                "\\n5. Abbreviations used: NA = not applicable \\n \\n ")
}
```

`r text`
